The Unicode Cookbook \linebreak for Linguists

### 2018-02-27 book/chapters/ipa_background.tex
closes issue #6
fixes bibliography; updates IPA reference

Originally, the IPA was a list of symbols with pronunciation examples
using words in different languages. In 1900 the symbols were first organized into
a chart and were given phonetic feature labels (e.g.\ for manner of
articulation among others ``plosives'', ``nasales'', ``fricatives'', for place of
articulation among others ``bronchiales'', ``laryngales'', ``labiales'' and for vowels
e.g.\ ``fermées'', ``mi-fermées'', ``mi-ouvertes'', ``ouvertes''). Throughout the last
century, the structure of the chart has changed with increases in phonetic
knowledge. Thus, similar to notational systems in other scientific disciplines,
the IPA reflects facts and theories of phonetic knowledge that have developed
over time. It is natural then that the IPA is modified occasionally to
accommodate scientific innovations and discoveries. In fact, updates are part of the
Association's mandate. These changes are captured in revisions to the IPA chart.
%\footnote{For a detailed history see:
%\url{https://en.wikipedia.org/wiki/History\_of\_the\_International\_Phonetic_Alphabet}}

@@

Several styles of transcription with the IPA are possible, as illustrated in the
\textit{Handbook}, and they are all valid.\footnote{For an illustration of
the differences, see the 29 languages and their transcriptions in the
\textit{Illustrations of the IPA} \citep[41--154]{IPA2007}.} Therefore, there are 
different but equivalent transcriptions, or as noted by \citet[64]{Ladefoged1990a}, 
``perhaps now that the Association has been explicit in its eclectic approach, outsiders to the
Association will no longer speak of \textit{the} IPA transcription of a given
phenomenon, as if there were only one approved style.'' Clearly not all
phoneticians agree, nor are they likely to ever completely agree, on all aspects of the
IPA or on transcription approaches and practices in general. As noted above, 
there have been several revisions in the IPA's long history, but the current version (2005) is
strikingly similar to the 1926 version, which shows the viability of the IPA as a
common standard for linguistic transcription.

@@

@@

The working group assigned each IPA symbol to a unique three-digit number, i.e.\
an IPA Number. Encoded in this number scheme implicitly includes information
about the status of each symbol (see below). The IPA numbers were listed with the
IPA symbols and they were also illustrated in IPA chart form (see
\cite[84]{EslingGaylord1993} or \cite[App. 2]{IPA2007}). The numbers were
assigned in linear order (e.g.\ [p] 101, [b] 102, [t] 103...) following the IPA
revision of 1989 and its update in 1996. Although the numbering scheme still 
exists, in practice is it superseded by the Unicode codification of symbols.

>>>>

Originally, the IPA was a list of symbols with pronunciation examples
using words in different languages. In 1900 the symbols were first organized into
a chart and were given phonetic feature labels (e.g.\ for manner of
articulation among others ``plosives'', ``nasales'', ``fricatives'', for place of
articulation among others ``bronchiales'', ``laryngales'', ``labiales'' and for vowels
e.g.\ ``fermées'', ``mi-fermées'', ``mi-ouvertes'', ``ouvertes''). Throughout the last
century, the structure of the chart has changed with increases in phonetic
knowledge. Thus, similar to notational systems in other scientific disciplines,
the IPA reflects facts and theories of phonetic knowledge that have developed
over time. It is natural then that the IPA is modified occasionally to
accommodate scientific innovations and discoveries. In fact, updates are part of the
Association's mandate. These changes are captured in revisions to the IPA chart.\footnote{For a detailed history see:
\url{https://en.wikipedia.org/wiki/History\_of\_the\_International\_Phonetic_Alphabet}}

@@

Several styles of transcription with the IPA are possible, as illustrated in the
\textit{Handbook}, and they are all valid.\footnote{For an illustration of
the differences, see the 29 languages and their transcriptions in the
\textit{Illustrations of the IPA} \citep[41--154]{IPA1999}.} Therefore, there are 
different but equivalent transcriptions, or as noted by \citet[64]{Ladefoged1990a}, 
``perhaps now that the Association has been explicit in its eclectic approach, outsiders to the
Association will no longer speak of \textit{the} IPA transcription of a given
phenomenon, as if there were only one approved style.'' Clearly not all
phoneticians agree, nor are they likely to ever completely agree, on all aspects of the
IPA or on transcription approaches and practices in general. As noted above, 
there have been several revisions in the IPA's long history, but the current version (2005) is
strikingly similar to the 1926 version, which shows the viability of the IPA as a
common standard for linguistic transcription.

@@

% TODO: cite Moran 2018

@@

The working group assigned each IPA symbol to a unique three-digit number, i.e.\
an IPA Number. Encoded in this number scheme implicitly includes information
about the status of each symbol (see below). The IPA numbers were listed with the
IPA symbols and they were also illustrated in IPA chart form (see
\cite[84]{EslingGaylord1993} or \cite[App. 2]{IPA1999}). The numbers were
assigned in linear order (e.g.\ [p] 101, [b] 102, [t] 103...) following the IPA
revision of 1989 and its update in 1996. Although the numbering scheme still 
exists, in practice is it superseded by the Unicode codification of symbols.

### 2018-02-27 book/chapters/practical_recommendations.tex
fixes bibliography; updates IPA reference

For general information about the IPA, the standard text is the \textit{Handbook of the International Phonetic Association: A Guide to the Use of the International Phonetic Alphabet} \citep{IPA2007}. The handbook describes in detail the principles and premises of the IPA, which we have summarized in Section \ref{IPApremises-principles}. The handbook also provides many examples of how to use the IPA. The Association also makes available information about itself online\footnote{\url{https://www.internationalphoneticassociation.org/}} and it provides the most current IPA charts.\footnote{\url{https://www.internationalphoneticassociation.org/content/ipa-chart}} Wikipedia also has a comprehensive article about the IPA.\footnote{\url{https://en.wikipedia.org/wiki/International_Phonetic_Alphabet}}

>>>>

For general information about the IPA, the standard text is the \textit{Handbook of the International Phonetic Association: A Guide to the Use of the International Phonetic Alphabet} \citep{IPA1999}. The handbook describes in detail the principles and premises of the IPA, which we have summarized in Section \ref{IPApremises-principles}. The handbook also provides many examples of how to use the IPA. The Association also makes available information about itself online\footnote{\url{https://www.internationalphoneticassociation.org/}} and it provides the most current IPA charts.\footnote{\url{https://www.internationalphoneticassociation.org/content/ipa-chart}} Wikipedia also has a comprehensive article about the IPA.\footnote{\url{https://en.wikipedia.org/wiki/International_Phonetic_Alphabet}}
